A Plasma Torus Around a Young Low-Mass Star

### 2025-03-20 paper/ms.tex
update

M dwarfs, stars with masses below about half that of the Sun, are the
only type of star to offer near-term prospects for detecting the
atmospheres of rocky exoplanets with surface water \cite{NAP26141}.
Investment with JWST has been proceeding accordingly
\cite{Redfield2024,TRAPPIST1JWSTCommunityInitiative2024}.  How an M
dwarf's evolution influences its planets---especially the retention of
their atmospheres--- has concurrently become a major theme in
exoplanet and stellar astrophysics.  Previous work has established
that most M dwarfs host close-in planets \cite{Dressing2015} that are
on average subject to long circumstellar disk lifetimes
\cite{Ribas2015}, to large doses of UV radiation \cite{France2013},
and to a high incidence of flares and coronal mass ejections
\cite{Gunther2020}.  However, despite excellent work in these areas,
the properties of the plasma and magnetospheric environments that
bathe young, close-in exoplanets have remained challenging to
quantify.

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.7\textwidth]{figures/f1.pdf}
  \captionsetup{labelformat=empty}
  \caption{{\bf Figure 1 (Movie):  TIC~141146667 is a complex periodic
  variable (CPV).}  The online movie, available
  \href{https://lgbouma.com/movies/TIC141146667_20250116.mp4}{here},
  covers a baseline of 5{,}784 cycles irregularly sampled over three
  years.  The TESS light curve is phased to the \periodhr\ hour period
  in groups of a few cycles per frame.  This is the period both of
  stellar rotation, and (we hypothesize) of corotating clumps of
  circumstellar material.  Raw data acquired with two minute sampling
  are in gray; black averages to 100 points per cycle.  Similar to
  other stars of this class, the sharp photometric features persist
  for tens to thousands of rotational cycles. }
  \label{fig:lc}
\end{figure}

@@

In this study, we present the first detection of corotating clumps of
cool plasma around a CPV.  We identified TIC~141146667 in previous
work \cite{Bouma2024} by searching the TESS two-minute data
\cite{Ricker2015} for stars showing at least three sharp, periodic
dips per cycle.  We selected the star for spectroscopy because its
brightness and rotation rate enable efficiently mapping out any line
profile variability.  We observed for five hours on 17~February~2024
(UT) using the High Resolution Echelle Spectrometer (HIRES
\cite{vogt_hires_1994}) on the Keck I 10m telescope.  TESS observed
the star over a four week interval from 30~January~2024 to
26~February~2024 with a duty cycle of 77\%.  During the spectral
observations, TESS was performing a data downlink; TESS data
collection resumed 12 hours (three rotation cycles) after the spectra
were acquired.  Extended Data Figure~\ref{fig:fulllc} shows the
detailed photometric behavior of the star before and after the exact
epoch of observation; the general photometric morphology is similar
before and after the data gap.

@@

\begin{figure}[!tp]
  \centering
  \includegraphics[width=0.99\textwidth]{figures/f2.png}
  \captionsetup{labelformat=empty}
  \caption{{\bf Figure 2 (Movie):}  
  Hydrogen emission from circumstellar plasma orbiting TIC 141146667.
  The online movie, available
  \href{https://lgbouma.com/movies/TIC141146667_sixpanel.mp4}{here},
  shows the spectral evolution over five hours.
  {\bf Panel a:} Average TESS light curve from 5 February 2024 to 26
  February 2024 (UT) folded on the \periodhr\ hour period.  Black
  points are phase-averaged; gray are the raw data.
  {\bf Panel b:} Keck/HIRES H$\alpha$ spectra acquired on UT 17
  February 2024.  The continuum is set to unity, and the darkest color
  is at twice the continuum to accentuate emission outside the line
  core ($|v/v_{\rm eq}|>1$, for $v_{\rm eq}$=130\,\kms).  While
  emission in the line core originates in the star's chromosphere,
  the sinusoidal emission features are most readily described by a
  warped plasma torus.
  {\bf Panel c:} Individual epochs of Panel b, visible in the
  online movie.  The dotted line shows the time-averaged spectrum,
  $f_{\langle t \rangle}$.
  {\bf Panel d:} As in Panel a, but overplotting the
  median-normalized H$\alpha$ light curve at $|v/v_{\rm eq}|<1$.
  {\bf Panel e:} As in Panel b, after subtracting the time-averaged
  spectrum. In addition to circumstellar emission, the line core shows
  absorption during the plasma clump transits.  The asymmetric stretch
  is set to match the dynamic range of the data.
  {\bf Panel f:} Individual epochs of Panel e, visible in the online
  movie.}
  \label{fig:spec}
\end{figure}

% photometry
Figure~\ref{fig:spec} shows the TESS and HIRES data from February
2024.  Over timescales of years, CPVs maintain a fixed period while
their photometric morphology evolves.  TIC~141146667 indeed evolved
relative to the February 2022 discovery data (see Figure~\ref{fig:lc}.
In February 2024, the average photometric signal showed a small
brightening over 45\% of the period, followed by a complex flux dip
spanning 55\% of the period.  This eclipse feature shows two to three
local photometric minima, and one to two local maxima.  Its W-shape is
suggestive of eclipse geometries seen in some forward models of warped
plasma tori \cite{Townsend2008}.

@@

% * keck/hires: behavior at v/veq<1
The behavior within the stellar H$\alpha$ line core, at $|\Delta v /
v_{\rm eq}|<1$, is more complex than outside it.  For stars of
this age and spectral type, one would expect emission in the line core
to be generated in the star's chromosphere
and then modulated by any occulting material capable of absorbing or
emitting H$\alpha$ photons.  In Figure~\ref{fig:spec}e, the behavior from
$\phi$=0.4-1.2 has a simple interpretation: from $\phi$=0.4-0.9, a hot
region first gradually crosses the stellar line profile, followed from
$\phi$=0.7-1.2 by the transit of a cool region.  Phases $\phi$$<$0.4
show a mix of similar events, though the time sampling is
sufficiently coarse that the interpretation is less clear.  A final
exercise to quantify the behavior in the line core is shown in
Figure~\ref{fig:spec}d, where $f_{\rm H\alpha\ core}$ denotes the
summed flux at $|\Delta v / v_{\rm eq}|<1$.  This panel shows that
changes in the line core flux are usually correlated with the
broadband variability, except at $\phi$=0.5, during the transit of the
3.9\,$R_\star$ clump and the occultation of the lower-velocity clump.

\begin{table*}
\scriptsize
\setlength{\tabcolsep}{2pt}
\centering
\caption{Selected System Parameters for TIC~141146667}
\label{tab:params}
\begin{tabular}{llcc}
\hline \hline
Parameter & Description & Value & Source\\
\hline 
%
$T_{\rm eff}$\dotfill                   & Effective Temperature (K) \hspace{9pt}\dotfill                 & 2972 $\pm$ 40    & 1 \\
%
$R_\star$\dotfill                       & Stellar radius ($R_\odot$)\dotfill                             & 0.42$\pm$0.02    & 1 \\
%
Age                                     & Stellar age range (Myr)\dotfill                                & 35-150           & 2 \\
%
$M_\star$\dotfill                       & Stellar mass ($M_\odot$)\dotfill                               & 0.22$\pm$0.02    & 3 \\
%
$\gamma$\dotfill                        & Systemic radial velocity (\kms)\dotfill                        & 0.61 $\pm$ 1.47  & 4 \\
%
Spec. Type\dotfill                      & Spectral Type\dotfill                                          & M5.5Ve           & 4 \\
%
$P_{\rm rot}$\dotfill                   & Rotation period (hr)\dotfill                                   & $3.930\pm 0.001$ & 5 \\
%
$v_{\rm eq}$\dotfill		                & Equatorial velocity \dotfill                                   &  130$\pm$4       & 6 \\
                                        & \hspace{3pt} ($2\pi R_\star/P_{\rm rot}$) (\kms)	             &                      \\
%
$v_{\rm eq}\sin{i_\star}$\dotfill		    & Projected rotational\dotfill                                   &  138$\pm$8       & 4 \\
                                        & \hspace{3pt} velocity (\kms)	                                 &                      \\
%
$i_\star$\dotfill                       & Stellar inclination\dotfill                                    & 	$>$63           & 4 \\
                                        & \hspace{3pt}  2$\sigma$ lower limit (deg)	                     &                      \\
%
$d$\dotfill                             & Distance (pc)\dotfill                                          & $57.54 \pm 0.09$ & 7 \\
%
$R_{\rm c}$\dotfill		                  & Keplerian corotation\dotfill                                   & $1.82 \pm 0.10$  & 6 \\
                                        & \hspace{3pt} radius ($R_\star$)	                               &                      \\
%
$a_1$\dotfill                           & Clump 1 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  2.1-2.7         & 4 \\
$a_2$\dotfill                           & Clump 2 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  3.9             & 4 \\
%
$\langle$EW$_{\rm H\alpha}$$\rangle$    & Time-averaged H$\alpha$ line core                              &  X.X             & Y \\ 
                                        & \hspace{3pt} equivalent width (\AA)	                           &                      \\
Range(EW$_{\rm H\alpha}$)$_1$           & H$\alpha$ equiv. width range                                   &  X.X             & Y \\ 
                                        & \hspace{3pt} from Clump 1 (\AA)	                               &                      \\
Range(EW$_{\rm H\alpha}$)$_2$           & H$\alpha$ equiv. width range                                   &  X.X             & Y \\ 
                                        & \hspace{3pt} from Clump 2 (\AA)	                               &                      \\
\hline
\end{tabular}
\begin{flushleft}
\footnotesize{ \textsc{NOTE}---
$^*$ Given only $v\sin i$ and $2\pi R_\star/P_{\rm rot}$, $\cos i=0.11^{+0.11}_{-0.08}$.
Provenances are:
1: SED fit \cite{Bouma2024}.
2: Pre-main sequence Gaia DR3 photometry, but lacks of lithium (see~\ref{subsec:stparams}).
3: PARSEC v1.2S \cite{Chen2014}.
4: Keck/HIRES.
5: TESS light curve.
6: Derived quantity.
7: Gaia DR3 geometric \cite{GaiaDR3}.
}
\end{flushleft}
\vspace{-0.5cm}
\end{table*}

@@

Spectra of magetically-active, rapidly rotating stars with a wide range
of masses have been known to exhibit both sinusoidal emission features
\cite{Donati2000,Townsend2005,Dunstone2006,Skelly2008} as well as
transient absorption features in their line cores
\cite{CollierCameron1989,CollierCameron1992,Cang2020} similar to
Figure~\ref{fig:spec}.  No such stars were previously known to show
complex light curves \cite{Bouma2024}.   One interpretation for
such spectroscopic variability comes from an analogy to quiescent solar
prominences, which are cool condensations of plasma in the solar corona
that can last days to weeks \cite{VialEngvold2015}.  These
condensations fall back to the solar surface because gravity is
stronger than any magnetic tension or centrifugal force capable of
sustaining them.  However for stars with magnetospheric radii $R_{\rm
m}$ that exceed their corotation radii $R_{\rm c}$, the effective
potential experienced by a charged particle can have a local minimum
outside $R_{\rm c}$, enabling the material to accumulate over much
longer timescales \cite{Petit2013,Daley-Yates2024}.  Generally
speaking, such material need neither transit, nor be optically thick.
For instance, the H$\alpha$ emission clump at 3.9\,$R_\star$ shows no 
photometric signal in the February 2024 photometry.

Our Keck/HIRES observations are the first reported time-series
spectra of a CPV, and they show that corotating circumstellar plasma clumps 
are present in at least one such star.
Based on the observed H$\alpha$ luminosity, the characteristic
densities and masses of the gaseous component of these clumps are
$n_{\rm H} \sim 10^{11}$-$10^{12}$\,cm$^{-3}$ and $M_{\rm gas} \sim 10^{17}$\,g
(see Section~\ref{subsec:gas}).
Any dust, if present, is independently constrained to have a total
mass $10^{15}\,{\rm g} < M_{\rm dust} < 10^{17}\,{\rm g}$ (see
Section~\ref{subsec:dust}).
This
observation rules out a ``starspot-only'' origin scenario for CPVs,
\cite{Koen2021} because such scenarios have no means of explaining
spectroscopic emission beyond the stellar disk.  Similarly, scenarios in
which the circumstellar material is made only of dust are also ruled
out.  While dust may be present, to explain the H$\alpha$ emission the
clumps must include plasma with a significant population of hydrogen
atoms in the $n$=3 excited state.  While this plasma is undoubtedly
sculpted by the star's magnetic field, it could plausibly originate from
three sites: the star, an old and undetected disk, or outgassing rocky
bodies.  This latter possibility would render CPVs as extrasolar analogs
of the Jupiter-Io plasma torus (e.g.~Ref.~\cite{Bagenal1981}).

The other potential analog for the CPVs are the $\sigma$~Ori~E
variables, a rare subset of B stars with radiatively-driven winds that
accumulate into warped plasma tori \cite{Townsend2005,Townsend2008}.
These tori tend to have dense antipodal accumulations of plasma sculpted
by tilted-dipole magnetic fields, and the transits of these clumps are
thought to produce the observed broadband photometric variability
through bound-free scattering \cite{Townsend2005} and Thomson scattering
\cite{Berry2022}.  For $\sigma$~Ori~E and almost all of its analogs, the
result is light curves that appear ``simple'', resembling those of
eclipsing binaries, and ``S-wave'' H$\alpha$ spectra similar to those
in Figure~\ref{fig:spec} \cite{Townsend2005,Townsend2008}.  The two known exceptions,
HD~37776 and HD~64740, show complex light curves resembling CPVs
\cite{Mikulasek2020,Bouma2024} and have spectropolarimetric magnetic
field maps indicating strong contributions from higher-order magnetic
moments \cite{Kochukhov2011,Shultz2018}.  There are two implications:
first, the complexity of CPVs may be a direct consequence of magnetic
fields with highly multipolar contributions.  Second, CPVs can be a
source of astrophysical false positives in photometric classification
of eclipsing binaries and transiting exoplanets around young pre-main
sequence M dwarfs \cite{Johns-Krull2016,Bouma2020}.

@@

The role of the star's magnetic field could be better understood through
new observations, and new theory.  From the theoretical perspective,
there is an urgent need for rigid-field (magneto)-hydrodynamic modeling
to go beyond previous work \cite{Townsend2005,Townsend2008} and to
explore the effects of non-dipolar field contributions.  In particular,
dynamo simulations of fully-convective M dwarfs have suggested that
global-scale mean fields might be confined to a single hemisphere
\cite{Brown2020}; such fields could yield accumulation surfaces quite
different from those that have previously been explored
\cite{Townsend2008}.  Observationally, spectropolarimetry has the
potential to assess both the field strength and topology.  An
independent probe could also be connected the recent work
\cite{Kaur2024} showing that CPVs are variable radio emitters with
emission components that can be not only persistent, but also
short-lived and highly polarized.  In particular, detecting radio
emission produced by an electron cyclotron maser instability
(e.g. \cite{Callingham2021}) would provide a measurement of the field
strength at the site of the emitting region.

@@

\begin{figure}[!b]
  \centering
  \includegraphics[width=0.99\textwidth]{figures/sf1.pdf}
  \caption{Photometric evolution of TIC 141146667 near the Keck/HIRES observation (green bar). 
  {\bf Panel a}: TESS simple aperture photometry. 
  The main data gaps were caused by scattered light from the Earth
  (BTJD 3356-3358.5) and Moon (BTJD 3359.5-3360.5).
  Raw two minute data are in gray; black time-averages ten minute
  sampling.
  {\bf Panels b,c}: Folded light curve before and after spectroscopy.
  Raw two minute data are in gray; black phase-averages to 100 points
  per cycle.
  During BTJD 3352-3356, a state switch occurred near BTJD 3353,
  in which the dip at $\phi$$\approx$0.8 disappeared.
  While the large eclipse feature remained present before and after
  the HIRES sequence, its photometric shape evolved during the data
  gap.
  }
  \label{fig:fulllc}
\end{figure}

{\it Photometry:} TESS observed TIC~141146667 ($T$=13.3) in Sectors 14,
15, 21, 41, 48, and 75.  Two-minute data were acquired during Sectors
41, 48 (TESS DDT039, PI: Kunimoto), and 75 (TESS Program G06030, PI:
Bouma).  The data from Sectors 14, 15, and 21 had a 30-minute cadence,
which smears sharp features over the $<$4\,hour period (see
\cite{Gunther2022}).  The field is not crowded:  the nearest known
star, TIC~141146666 ($T$=14.5), is 25$''$ from TIC~141146667 and is
photometrically stable in the pixel-level TESS data.

@@

{\it Viewing Orientation}---We fitted the rotational broadening of the
K~I (7700\,\AA) resonance line using the broadening kernel suggested by
\cite{Gray2008}; taking the mean and standard deviation of the resulting
value over all epochs yielded $v_{\rm eq} \sin i$=138$\pm$8\,\kms,
consistent with the visual line broadening $\Delta
\lambda$$\approx$3\,\AA.  The star's equatorial velocity $v_{\rm eq}$
based on its apparent size and rotation period is 130$\pm$4\,\kms.
While this suggests that the viewing orientation could be nearly
edge-on, the formal constraint is rather weak, with $i$$>$63$^\circ$ at
2$\sigma$ (2.5$^{\rm th}$ percentile of the inclination posterior).

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.7\textwidth]{figures/sf4.pdf}
  \caption{
    Spectral energy distribution of broadband photometric magnitudes
    plotted over the best-fit BT-Settl stellar atmosphere model
    \cite{Allard2012}.  This plot was made from an adaptation of \texttt{astroARIADNE}
    \cite{Vines2022}.  The photometry extends from the Gaia DR3 blue
    passband to WISE W3;  the W4 passband (22 $\mu$m) did not yield
    a confident detection.  This fit yields the star's temperature and
    size.  The lack of excess infrared flux relative to the photospheric
    model sets an upper limit on emission from circumstellar dust.
    }
  \label{fig:sed}
\end{figure}

>>>>

M dwarfs, stars with masses below about half that of the Sun, are the
only type of star to offer near-term prospects for detecting the
atmospheres of rocky exoplanets with surface water \cite{NAP26141}.
Community investment with JWST has proceeded accordingly
\cite{Redfield2024,TRAPPIST1JWSTCommunityInitiative2024}.  How an M
dwarf's evolution influences its planets---especially the retention of
their atmospheres--- has in turn become a major theme in
exoplanet and stellar astrophysics.  Previous work has established
that most M dwarfs host close-in planets \cite{Dressing2015} that are
on average subject to long circumstellar disk lifetimes
\cite{Ribas2015}, to large doses of UV radiation \cite{France2013},
and to a high incidence of flares and coronal mass ejections
\cite{Gunther2020}.  However, despite the excellent work in these areas,
the properties of the plasma and magnetospheric environments that
bathe young, close-in exoplanets have remained challenging to
quantify.

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.7\textwidth]{figures/f1.pdf}
  \captionsetup{labelformat=empty}
  \caption{{\bf Figure 1 (Movie):  TIC~141146667 is a complex periodic
  variable (CPV).}  The online movie,
  \href{https://lgbouma.com/movies/TIC141146667_20250116.mp4}{available here},
  covers a baseline of 5{,}784 cycles irregularly sampled over three
  years.  The TESS light curve is phased to the \periodhr\ hour period
  in groups of a few cycles per frame.  This is the period both of
  stellar rotation, and (we hypothesize) of corotating clumps of
  circumstellar material.  Raw data acquired with two minute sampling
  are in gray; black averages to 100 points per cycle.  Similar to
  other stars of this class, the sharp photometric features persist
  for tens to thousands of rotational cycles. }
  \label{fig:lc}
\end{figure}

@@

In this study, we present the first spectroscopic detection of corotating clumps of
cool plasma around a CPV.  We identified TIC~141146667 in previous
work \cite{Bouma2024} by searching the TESS two-minute data
\cite{Ricker2015} for stars showing highly structured, periodic light
curves.  We selected the star for spectroscopy because its
brightness and rotation rate enable efficiently mapping out line
profile variability.  We observed it for five hours on 17~February~2024
(UT) using the High Resolution Echelle Spectrometer (HIRES
\cite{vogt_hires_1994}) on the Keck I 10m telescope.  TESS observed
the star over a four week interval from 30~January~2024 to
26~February~2024 with a duty cycle of 77\%.  During the HIRES
observations, TESS was performing a data downlink; TESS data
collection resumed 12 hours (three rotation cycles) after the spectra
were acquired.  Extended Data Figure~\ref{fig:fulllc} shows the
detailed photometric behavior of the star near the
epoch of observation; the general photometric morphology is similar
before and after the data gap.

@@

\begin{figure}[!tp]
  \centering
  \includegraphics[width=0.99\textwidth]{figures/f2.pdf}
  \captionsetup{labelformat=empty}
  \caption{{\bf Figure 2 (Movie):}  
  Emission from circumstellar plasma orbiting TIC 141146667.
  The online movie,
  \href{https://lgbouma.com/movies/TIC141146667_sixpanel.mp4}{available here},
  shows the spectral evolution over five hours.
  {\bf Panel a:} Average TESS light curve from 5 February 2024 to 26
  February 2024 (UT) folded on the \periodhr\ hour period.  Black
  points are phase-averaged; gray are the raw data.
  {\bf Panel b:} Keck/HIRES H$\alpha$ spectra acquired on 17
  February 2024.  The continuum is set to unity, and the darkest color
  is at twice the continuum to accentuate emission outside the line
  core ($|v/v_{\rm eq}|>1$, for $v_{\rm eq}$=130\,\kms).  While
  emission in the line core originates in the star's chromosphere,
  the sinusoidal emission features are most readily described by a
  warped plasma torus.
  {\bf Panel c:} Individual epochs of Panel b, visible in the
  online movie.  The dotted line shows the time-averaged spectrum,
  $f_{\langle t \rangle}$.
  {\bf Panel d:} As in Panel a, but overplotting the
  median-normalized H$\alpha$ light curve at $|v/v_{\rm eq}|<1$.
  {\bf Panel e:} As in Panel b, after subtracting the time-averaged
  spectrum.  The line core shows H$\alpha$ excesses and decrements
  advancing from the blue to red wings.
  The asymmetric color stretch
  is set to match the dynamic range of the data.
  {\bf Panel f:} Individual epochs of Panel e, visible in the online
  movie.}
  \label{fig:spec}
\end{figure}

% photometry
Figure~\ref{fig:spec} compiles the TESS and HIRES data from February
2024.  Over timescales of years, CPVs maintain a fixed period while
their photometric morphology evolves.  TIC~141146667 indeed evolved
relative to the February 2022 discovery data (see Figure~\ref{fig:lc}).
In February 2024, the average photometric signal showed a small
brightening over 45\% of the period, followed by a complex flux dip
spanning 55\% of the period.  This eclipse feature shows two to three
local photometric minima, and one to two local maxima.  Its W-shape is
suggestive of eclipse geometries seen in certain forward models of warped
plasma tori \cite{Townsend2008}.

@@

% * keck/hires: behavior at v/veq<1
Within the stellar H$\alpha$ line core, at $|\Delta v /
v_{\rm eq}|<1$, we observe a superposition of emission from the star's
chromosphere,
modulated by any occulting material capable of absorbing or
emitting H$\alpha$ photons.  In Figure~\ref{fig:spec}e, the data from
$\phi$=0.4-1.2 have a simple interpretation: from $\phi$=0.4-0.9, a hot
(emitting) region first gradually crosses the stellar line profile, followed from
$\phi$=0.7-1.2 by the transit of a cool (absorbing) region. Phases $\phi$$<$0.4
show a mix of similar events, though the time sampling is
sufficiently coarse that the interpretation is less clear.  A final
exercise to quantify the behavior in the line core is shown in
Figure~\ref{fig:spec}d, where $f_{\rm H\alpha\ core}$ denotes the
summed flux at $|\Delta v / v_{\rm eq}|<1$.  This panel shows that
changes in the line core flux are correlated with the
broadband variability over most of the light curve, except at
$\phi$$\approx$0.5, during the transit of the 3.9\,$R_\star$ clump and
the occultation of the lower-velocity clump.

\begin{table*}
\small
\setlength{\tabcolsep}{2pt}
\centering
\caption{Selected System Parameters for TIC~141146667}
\label{tab:params}
\begin{tabular}{llcc}
\hline \hline
Parameter & Description & Value & Source\\
\hline 
%
$T_{\rm eff}$\dotfill                   & Effective Temperature (K) \hspace{9pt}\dotfill                 & 2972 $\pm$ 40    & 1 \\
%
$R_\star$\dotfill                       & Stellar radius ($R_\odot$)\dotfill                             & 0.42$\pm$0.02    & 1 \\
%
Age                                     & Stellar age range (Myr)\dotfill                                & 35-150           & 2 \\
%
$M_\star$\dotfill                       & Stellar mass ($M_\odot$)\dotfill                               & 0.22$\pm$0.02    & 3 \\
%
$\gamma$\dotfill                        & Systemic radial velocity (\kms)\dotfill                        & 0.61 $\pm$ 1.47  & 4 \\
%
Spec. Type\dotfill                      & Spectral Type\dotfill                                          & M5.5Ve           & 4 \\
%
$P_{\rm rot}$\dotfill                   & Rotation period (hr)\dotfill                                   & $3.930\pm 0.001$ & 5 \\
%
$v_{\rm eq}$\dotfill		                & Equatorial velocity \dotfill                                   &  130$\pm$4       & 6 \\
                                        & \hspace{3pt} ($2\pi R_\star/P_{\rm rot}$) (\kms)	             &                      \\
%
$v_{\rm eq}\sin{i_\star}$\dotfill		    & Projected rotational\dotfill                                   &  138$\pm$8       & 4 \\
                                        & \hspace{3pt} velocity (\kms)	                                 &                      \\
%
$i_\star$\dotfill                       & Stellar inclination\dotfill                                    & 	$>$63           & 4 \\
                                        & \hspace{3pt}  2$\sigma$ lower limit (deg)	                     &                      \\
%
$d$\dotfill                             & Distance (pc)\dotfill                                          & $57.54 \pm 0.09$ & 7 \\
%
$R_{\rm c}$\dotfill		                  & Keplerian corotation\dotfill                                   & $1.82 \pm 0.10$  & 6 \\
                                        & \hspace{3pt} radius ($R_\star$)	                               &                      \\
%
$a_1$\dotfill                           & Clump 1 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  2.1-2.7         & 4 \\
$a_2$\dotfill                           & Clump 2 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  3.9             & 4 \\
%
$\langle$EW$_{\rm H\alpha}$$\rangle$    & Time-averaged H$\alpha$ line core                              &  7.2 $\pm$ 0.2   & 4 \\ 
                                        & \hspace{3pt} equivalent width (\AA)	                           &                      \\
\hline
\end{tabular}
\begin{flushleft}
\footnotesize{ \textsc{NOTE}---
Provenances are:
1: SED fit \cite{Bouma2024}.
2: Gaia DR3 photometry shows the star is on the pre-main sequence,
   while the spectrum lacks lithium (see~\ref{subsec:stparams}).
3: PARSEC v1.2S \cite{Chen2014}.
4: Keck/HIRES.
5: TESS light curve.
6: Derived quantity.
7: Gaia DR3 geometric \cite{GaiaDR3}.
}
\end{flushleft}
\vspace{-0.5cm}
\end{table*}

@@

Spectra of magetically-active, rapidly rotating stars with a wide range
of masses have been known to exhibit both sinusoidal emission features
\cite{Donati2000,Townsend2005,Dunstone2006,Skelly2008} as well as
transient absorption features
\cite{CollierCameron1989,CollierCameron1992,Cang2020} similar to
Figure~\ref{fig:spec}.  No such stars were previously known to show
complex light curves \cite{Bouma2024}.   One interpretation for
such spectroscopic variability comes from an analogy to quiescent solar
prominences, which are cool condensations of plasma in the solar corona
that can last days to weeks \cite{VialEngvold2015}.  These
condensations fall back to the solar surface because gravity is
stronger than any magnetic tension or centrifugal force capable of
sustaining them.  However for stars with magnetospheric radii $R_{\rm
m}$ that exceed their corotation radii $R_{\rm c}$, the effective
potential experienced by a charged particle can have a local minimum
outside $R_{\rm c}$, enabling the material to accumulate over much
longer timescales \cite{Petit2013,Daley-Yates2024}.  Generally
speaking, such material need neither transit, nor be optically thick.
For instance, the H$\alpha$ emission clump at 3.9\,$R_\star$ shows no 
photometric signal in the February 2024 photometry.

Our Keck/HIRES observations are the first reported time-series
spectra of a CPV, and they show that corotating circumstellar plasma clumps 
are present in at least one such star.
This
observation rules out a ``starspot-only'' origin scenario for CPVs,
\cite{Koen2021} because such scenarios have no means of explaining
spectroscopic emission beyond the stellar disk.  Similarly, scenarios in
which the circumstellar material is made only of dust are also ruled
out.  While dust may be present, to explain the H$\alpha$ emission the
clumps must include plasma with a significant population of hydrogen
atoms in the $n$=3 excited state.  
More quantitatively, the observed H$\alpha$ luminosity suggests
characteristic densities and masses for the gaseous component of these
clumps of $n_{\rm H} \sim 10^{11}$-$10^{12}$\,cm$^{-3}$ and $M_{\rm
gas} \sim 10^{17}$\,g (see Section~\ref{subsec:gas}).
Dust, if present, is independently constrained to have a total
mass $10^{15}\,{\rm g} < M_{\rm dust} < 10^{17}\,{\rm g}$ (see
Section~\ref{subsec:dust}).
This circumstellar material could plausibly originate from
three sites: the star, an old and undetected disk, or outgassing rocky
bodies.  This latter possibility would render CPVs as extrasolar analogs
of the Jupiter-Io plasma torus (e.g.~Ref.~\cite{Bagenal1981}).

The other potential analog for the CPVs are the $\sigma$~Ori~E
variables, a rare subset of B stars with radiatively-driven winds that
accumulate into warped plasma tori \cite{Townsend2005,Townsend2008}.
These tori tend to have dense antipodal accumulations of plasma sculpted
by tilted-dipole magnetic fields, and the transits of these clumps are
thought to produce the observed broadband photometric variability
through bound-free scattering \cite{Townsend2005} and Thomson scattering
\cite{Berry2022}.  For $\sigma$~Ori~E and almost all of its analogs, the
result is light curves that appear ``simple'', resembling those of
eclipsing binaries, and time-dynamic H$\alpha$ spectra similar to those
in Figure~\ref{fig:spec} \cite{Townsend2005,Townsend2008}.  The two known exceptions,
HD~37776 and HD~64740, show complex light curves resembling CPVs
\cite{Mikulasek2020,Bouma2024} and have spectropolarimetric magnetic
field maps indicating strong contributions from higher-order magnetic
moments \cite{Kochukhov2011,Shultz2018}.  There are two implications:
first, the complexity of CPVs may be a direct consequence of magnetic
fields with highly multipolar contributions.  Second, CPVs can be a
source of astrophysical false positives in photometric classification
of eclipsing binaries and transiting exoplanets around young pre-main
sequence M dwarfs \cite{Johns-Krull2016,Bouma2020}.

@@

The role of the star's magnetic field could be better understood through
new observations, and new theory.  From the theoretical perspective,
there is an urgent need for rigid-field (magneto)-hydrodynamic modeling
to go beyond previous work \cite{Townsend2005,Townsend2008} and to
explore the effects of non-dipolar field contributions.  In particular,
dynamo simulations of fully-convective M dwarfs have suggested that
global-scale mean fields might be confined to a single hemisphere
\cite{Brown2020}; such fields could yield accumulation surfaces quite
different from those that have previously been explored
\cite{Townsend2008}.  Observationally, spectropolarimetry has the
potential to assess both the field strength and topology.  An
independent probe could also be connected the recent work
\cite{Kaur2024} showing that CPVs are variable radio emitters with
emission components that can be not only persistent, but also
short-lived and highly polarized.  In particular, detecting radio
emission produced by an electron cyclotron maser instability
(e.g. \cite{Callingham2021}) would provide a measurement of the field
strength at the site of the emitting region, further clarifying the
magnetic field's structure.

@@

\begin{figure}[!b]
  \centering
  \includegraphics[width=0.99\textwidth]{figures/sf1.pdf}
  \caption{Photometric evolution of TIC 141146667 near the Keck/HIRES observation (green bar). 
  {\bf Panel a}: TESS simple aperture photometry. 
  The main data gaps were caused by scattered light from the Earth
  (BTJD 3356-3358.5) and Moon (BTJD 3359.5-3360.5).
  Raw two minute data are in gray; black time-averages to ten minute
  sampling.
  {\bf Panels b,c}: Folded light curve before and after spectroscopy.
  Raw two minute data are in gray; black phase-averages to 100 points
  per cycle.
  During BTJD 3352-3356, a state switch occurred near BTJD 3353,
  in which the dip at $\phi$$\approx$0.8 disappeared.
  While the large eclipse feature was present both before and after
  the HIRES sequence, its photometric shape evolved during the data
  gap.
  }
  \label{fig:fulllc}
\end{figure}

{\it Photometry:} TESS observed TIC~141146667 ($T$=13.3) in Sectors 14,
15, 21, 41, 48, and 75.  Two-minute data were acquired during Sectors
41, 48 (TESS DDT039, PI: Kunimoto), and 75 (TESS Program G06030, PI:
Bouma).  The data from Sectors 14, 15, and 21 had 30-minute cadence,
which smears sharp features over the $<$4\,hour period (see
\cite{Gunther2022}).  The field is not crowded:  the nearest known
star, TIC~141146666 ($T$=14.5), is 25$''$ from TIC~141146667 and is
photometrically stable in the pixel-level TESS data.

@@

{\it Viewing Orientation}---We fitted the rotational broadening of the
K~I (7700\,\AA) resonance line using the kernel suggested by
\cite{Gray2008}; taking the mean and standard deviation of the resulting
value over all epochs yielded $v_{\rm eq} \sin i$=138$\pm$8\,\kms,
consistent with the visual line broadening $\Delta
\lambda$$\approx$3\,\AA.  The star's equatorial velocity $v_{\rm eq}$
based on its apparent size and rotation period is 130$\pm$4\,\kms.
While this suggests that the viewing orientation could be nearly
edge-on, the formal constraint is rather weak, with $i$$>$63$^\circ$ at
2$\sigma$ (2.5$^{\rm th}$ percentile of the inclination posterior).

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.7\textwidth]{figures/sf4.pdf}
  \caption{
    Spectral energy distribution of broadband photometric magnitudes
    (filled markers)
    plotted over the best-fit BT-Settl stellar atmosphere model
    \cite{Allard2012} and the associated photometric predictions
    (empty markers).  This plot was made from an adaptation of
    \texttt{astroARIADNE} \cite{Vines2022}.  The photometry extends
    from the Gaia DR3 blue passband to WISE W3;  the W4 passband (22
    $\mu$m) did not yield a confident detection.  This fit yields the
    star's temperature and size.  The lack of excess infrared flux
    relative to the photospheric model sets an upper limit on emission
    from circumstellar dust.
    }
  \label{fig:sed}
\end{figure}
